---
title: "GW14 QC"
output:
  html_document:
    df_print: paged
---

```{r load-libraries}
library(Seurat)
library(readr)
library(ggplot2)
library(scales)
library(rcartocolor)
library(viridis)
library(tidyr)
library(dplyr)
library(janitor)
```

```{r import-data, cache=TRUE}
load("../ubuntu/gw14.RData")
gw14.2 <- CreateSeuratObject(raw.data = gw14@raw.data, min.cells = 0, min.genes = 0, project="GW14")
meta.14 <- read_delim("colnames_GW14_080818_fixed.txt", col_names = c("cell.name", "brain.region"), delim="\t")

ribogenes <- read_table("ribogenes.txt", col_names = F)
ribogenes <- ribogenes$X1

rm(gw14)
```

### Declare multiplot function.
```{r multiplot-fn}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

### Get HEX codes for 'Vivid' color palette.
```{r}
carto_pal(12, "Vivid")
display_carto_pal(12, "Vivid")
# c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#CC61B0" ,"#DAA51B", "#2F8AC4", "#764E9F")
```

### Set the brain region of cells as their original identity.
```{r set-ident}
gw14.2@meta.data$orig.ident <- factor(meta.14$brain.region)

gw14.2 <- SetIdent(gw14.2, cells.use = WhichCells(gw14.2), ident.use = meta.14$brain.region)
```

```{r}
summary(gw14.2@meta.data)
```

### Number and percentage of cells per brain region.
```{r }
gw14.2@meta.data %>% tabyl(orig.ident) %>% adorn_totals(where="row") %>% arrange(desc(n)) 
```

### Seurat's ViolinPlot plotting by area.
```{r eval=FALSE}
violin.14 <- VlnPlot(gw14.2, features.plot = c("nGene", "nUMI"), do.sort = T, point.size.use = .5, ident.include = as.character(statsByArea$orig.ident))
```

### Calculate % of mitochondrial genes in each cell.
```{r calc-mito}
mito.genes <- grep(pattern="^MT-", x=rownames(gw14.2@data), value = TRUE)
percent.mito <- colSums(gw14.2@raw.data[mito.genes, ])/colSums(gw14.2@raw.data)

gw14.2 <- AddMetaData(object = gw14.2, metadata = percent.mito, col.name = "percent.mito")
```

### Calculate % of ribosomal genes in each cell.
```{r calc-ribo}
ribo <- gw14.2@data[ribogenes, ]
percent.ribo <- colSums(ribo)/colSums(gw14.2@raw.data)

gw14.2 <- AddMetaData(object = gw14.2, metadata = percent.ribo, col.name = "percent.ribo")
```

### Calculate mean UMIs, genes, ribosomal and mitochondrial % for each brain region.
```{r calc-stats-by-area}
statsByArea <- gw14.2@meta.data %>% dplyr::group_by(orig.ident) %>% 
  summarise(mean.nGene = mean(nGene), mean.nUMI =mean(nUMI), mean.pctmito = mean(percent.mito), mean.pctribo= mean(percent.ribo))

statsByArea
```


## PLOTS

### 1. Violin plot: Genes per cell, split by brain region.
```{r p1}
gw14.2@meta.data$orig.ident <- ordered(gw14.2@meta.data$orig.ident, 
                                       levels = (statsByArea %>% arrange(mean.nGene))$orig.ident)

p1 <- ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=orig.ident), alpha=0.2, size=0.25) + 
  geom_violin(aes(orig.ident, nGene), colour="grey60", scale = "width", alpha=0.3) +
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  scale_y_continuous(breaks=seq(0,8000, by=1000)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Genes per Cell",
       x="", 
       y="Genes per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p1
```


### 2. Violin plot: Genes per cell [zoom-in].
```{r p2}
p2 <- ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=orig.ident), alpha=0.2, size=0.25) + 
  geom_violin(aes(orig.ident, nGene), colour="grey60", scale = "width", alpha=0.3) +
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  scale_y_continuous(breaks=seq(0,4000, by=500), limits = c(0,4000)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Genes per Cell (Zoom)",
       x="", 
       y="Genes per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)
  
p2
```



### 3. Violin plot: UMIs per cell, split by brain region.
```{r p3}
p3 <- 
  ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nUMI, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60", colour="grey30", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  scale_y_continuous(breaks=seq(0,60000, by=10000)) +
  labs(title="GW14: UMIs per Cell",
       x="", 
       y="UMIs per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p3
```

### 4. Violin plot: UMIs per cell [Zoom-in].
```{r p4}
p4 <- 
  ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nUMI, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60",  colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, 10000, by=1000), limits=c(0, 10000)) +
  labs(title="GW14: UMIs per Cell (Zoom)",
       x="", 
       y="UMIs per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p4
```

### 5. Violin plot: Genes per cell + color by nUMI.
```{r p5}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(nUMI)

p5 <- ggplot(gw14.2@meta.data %>% filter(nGene < 3500)) + 
  geom_jitter(aes(orig.ident, nGene, colour=nUMI), alpha=0.6, size=0.25) + 
  scale_color_viridis(option="magma", begin = .1, end = .95) +
  geom_violin(aes(orig.ident, nGene), colour="grey60",  colour="grey60", alpha=0.1, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.9,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 12),
        # legend.title.align = 0.9
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, 40000, by=500), limits=c(0, 3500)) +
  labs(title="GW14: Genes per Cell / UMIs per Cell",
       x="", 
       y="Genes Detected",
       color= "")

p5
```

### 6. Violin plot: UMIs per cell + color by nGenes.
```{r p6}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(nGene)

p6 <- ggplot(gw14.2@meta.data %>% filter(nUMI < 10000)) + 
  geom_jitter(aes(orig.ident, nUMI, colour=nGene), alpha=0.75, size=0.25) + 
  scale_color_viridis(option="magma", begin = .1, end = .9) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60", alpha=0.1, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.9,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 12),
        # legend.title.align = 0.9
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, 10000, by=1000), limits=c(0, 10000)) +
  labs(title="GW14: UMIs per Cell / Genes per Cell",
       x="", 
       y="UMIs Detected",
       color= "")

p6
```

### 7. % Ribosomal, split by brain region.
```{r p7}
p7 <- 
ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, percent.ribo, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  geom_violin(aes(orig.ident, percent.ribo), colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust=0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  # scale_y_continuous(breaks=seq(0,40000, by=5000)) +
  # ylim(0,40000) +
  labs(title="GW14: % Ribosomal Genes per Cell",
       x="", 
       y="% Ribosomal Genes"
      ) +
  guides(fill=FALSE, colour=FALSE)

p7
```

### 8. % Mitochondrial, split by brain region.
```{r p8}
p8 <- 
  ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, percent.mito, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F")) +
  geom_violin(aes(orig.ident, percent.mito), colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust=0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  # scale_y_continuous(breaks=seq(0,0.1, by=0.01), limits = c(0, 0.1)) +
  
  labs(title="GW14: % Mitochondrial Genes per Cell",
       x="", 
       y="% Mitochondrial Genes"
      ) +
  guides(fill=FALSE, colour=FALSE)

p8
```


### 9. Violin plot: Genes per cell; color by % ribosomal genes.
```{r p9}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(percent.ribo)

ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.ribo), alpha=0.6, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey40", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .9) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  scale_y_continuous(breaks=seq(0,8000, by=500)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Genes Detected per Cell",
       x="Area", 
       y="nGene"
      )
```

### 10. Violin plot: Genes per cell; color by % ribosomal genes [zoom-in]
```{r p10}
p10 <- ggplot(gw14.2@meta.data %>% filter(nGene < 4000)) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.ribo), alpha=0.8, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey40", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .1, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        # legend.title.align = 0.9
        ) +
  scale_y_continuous(breaks=seq(0,3500, by=500), limits = c(0,3500)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Genes per Cell / % Ribosomal Genes\n(zoom-in)",
       x="", 
       y="Genes per Cell",
       color="% ribosomal"
      )

p10
```


### 11. Violin Plot: Genes per cell, color by % mitochondrial genes
```{r p11}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(percent.mito)

p11 <- ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.mito), alpha=0.7, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .9) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1)
        ) +
  scale_y_continuous(breaks=seq(0,8000, by=500)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Number of Genes Detected per Cell",
       x="Area", 
       y="nGene"
      )

p11
```

### 12. Violin Plot: Genes per cell, color by % mitochondrial genes [zoom-in]
```{r p12}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(percent.mito)

p12 <- ggplot(gw14.2@meta.data %>% filter(nGene < 4000)) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.mito), alpha=0.8, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        # legend.title.align = 0.9
        ) +
  scale_y_continuous(breaks=seq(0,3500, by=500), limits = c(0,3500)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title="GW14: Genes per Cell / % Mitochondrial Genes",
       x="", 
       y="Genes per Cell",
       color="% mito"
      )

p12
```

### 13. Violin Plot: % Ribosomal, color by % mitochondrial [zoom-in]
```{r p13}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(percent.mito)

p13 <- 
ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, percent.ribo, colour=percent.mito), alpha=0.7, size=0.25) + 
  geom_violin(aes(orig.ident, percent.ribo), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .10, end = .90) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.87,.95),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        # legend.title.align = 0.9
        ) +
  # scale_y_continuous(breaks=seq(0,40000, by=5000)) +
  # ylim(0,40000) +
  labs(title="GW14: % Ribosomal Genes / % Mitochondrial Genes per Cell",
       x="", 
       y="% Ribosomal Genes",
       color="% mito"
      )

p13
```


### 14. Violin Plot: % Mitochondrial, color by % ribosomal [zoom-in]
```{r p14}
gw14.2@meta.data <- gw14.2@meta.data %>% arrange(percent.ribo)

p14 <- 
ggplot(gw14.2@meta.data) + 
  geom_jitter(aes(orig.ident, percent.mito, colour=percent.ribo), alpha=0.7, size=0.20) + 
  geom_violin(aes(orig.ident, percent.mito), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .10, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,.93),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        # legend.title.align = 0.9
        ) +
  # scale_y_continuous(breaks=seq(0,40000, by=5000)) +
  # ylim(0,40000) +
  labs(title="GW14: % Mitochondrial Genes / % Ribosomal Genes per Cell",
       x="", 
       y="% Mitochondrial Genes",
       color = "% ribosomal"
      )

p14
```

```{r print-plots}
png("gw14_QC_1.png", width=20, height=20, units="in", res=300)

multiplot(p1, p3, p7, p8, p2, p4, p10, p12, p5, p6, p13, p14, cols=3)
#multiplot( p10, p12, p13, p14, cols=3)

dev.off()
```

