---
title: "Clustering Pipeline with Seurat, PCA and Jaccard Distances - from Aparna"
date: "Last modified: Sep 20, 2018"
author: "Modified by: Carmen"
output: html_notebook
---

## 0. Setup
```{r}
#!/usr/bin/Rscript

# Obtain the ID being accessed from the array
jobid = as.integer(Sys.getenv("PBS_ARRAYID"))

# Expect command line args at the end.
args = commandArgs(trailingOnly = TRUE)

  age <- as.numeric(args[1])
  
  # age <- "gw22"
  
  perplexity <- as.numeric(args[2])
```

### 0.0 Declare functions.
```{r}

# 1
pushbullet.start <- function(message){

  RPushbullet::pbPost("note",
    title = paste0(age, ": Started"), body = message,
    debug = F, devices = c("Chrome"))
}

    pushbullet.start(message=paste0("R Script: Clustering"))

# 2
options(error = function() {
  library(RPushbullet)
  pbPost("note", title=paste("Error:", age), body=geterrmessage())
})

# 3
pushbullet.toc <- function(){

  toc(log = TRUE, quiet = TRUE)
  message <- tic.log(format = TRUE)[[1]]
  RPushbullet::pbPost("note", title=paste0(age, ": Finished"), body=message,
    devices = c("Chrome"))
  tic.clearlog()

}

# sort(sapply(ls(), function(x) format(object.size(get(x)), unit = 'auto')))

```

### 0.1 Load libraries.
```{r}
library(Seurat)
library(gmodels)
library(dplyr)
library(Matrix)
library(RANN)
library(igraph)
library(tictoc)
source("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/code/doFastPCA.R")
```

### 0.2 Load original Seurat object.
```{r}
# Trial datasets:
# gw20_34cortex.RData
# gw22.RData

tic("Loading object")

  load(paste0("/kriegsteinlab/data1/aparna/homefiles/", age, ".RData"))

pushbullet.toc()
```

## 1. Normalize and scale raw data

### 1.1 Create Seurat object:
```{r}
# Object initialization for Seurat v2.3:

tic("Initialize Seurat object")

    assign("seurat.object", CreateSeuratObject(display.progress = F,
                                              raw.data = get(age)@raw.data, 
                                              min.cells = 100, min.genes = 500,
                                              project=age)
          )

    rm(list=age)
    
pushbullet.toc()
```

### 1.2 Log-normalize data
```{r}
tic("Log-normalize data")

seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", 
                                 scale.factor = 1e4, display.progress = F)

pushbullet.toc()
```

### 1.3 Remove high mitochondrial % cells:
```{r}
tic("Remove high mitochondrial % cells")
  
  mito.genes <- grep("^MT-", rownames(seurat.object@data), value = T)
  
  percent.mito <- colSums(expm1(seurat.object@data[mito.genes,])) / colSums(expm1(seurat.object@data))
  
  seurat.object <- AddMetaData(seurat.object, percent.mito, "percent.mito")
  
  seurat.object <- SubsetData(seurat.object, subset.name = "percent.mito", accept.high = 0.10)
  
pushbullet.toc()
```

### 1.4 Scale

#### 1.4a Scale only (without nUMI regression):
```{r}
tic("Scale data")

seurat.object.scale <- ScaleData(seurat.object, do.scale = T, do.center = T, display.progress = F)

path.scaled.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "_seuratObject.scaleOnly.Rds")
saveRDS(seurat.object.scale, file=path.scaled.object)

pushbullet.toc()
```

#### 1.4b Scale with nUMI regression and store in a new object:
```{r}
tic("Scale data and regress nUMI")

seurat.object.scale.regress <- ScaleData(seurat.object, vars.to.regress = "nUMI", 
                                     do.scale = T, do.center = T, display.progress = F)

path.scaled.regressed.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "SeuratObject.scale.nUMI.Regress.Rds")
saveRDS(seurat.object.scale.regress, file=path.scaled.regressed.object)

pushbullet.toc()

seurat.objects <- list(scaled=seurat.object.scale, scaled.regressed=seurat.object.scale.regress)

rm(seurat.object.scale, seurat.object.scale.regress, seurat.object)

```

## 2. Run PCA

### 2.1 Run PCA on both objects:

```{r}
# Read-in scaled objects:

# path.scaled.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "_seuratObject.scaleOnly.Rds")
# path.scaled.regressed.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "SeuratObject.scale.nUMI.Regress.Rds")
# seurat.objects <- list(scaled=readRDS(path.scaled.object), scaled.regressed=readRDS(path.scaled.regressed.object))

selectObject <- function(object){
  
  path.scaled.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "_seuratObject.scaleOnly.Rds")
  path.scaled.regressed.object <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/data/scaledObjects/", age, "SeuratObject.scale.nUMI.Regress.Rds")
  
                           if(object == "scaled") {
                             scaling <<- "_scaleOnly"
                             pca.result <<- "pca.scaled"
                             object.path <<- path.scaled.object
                            } 
  
                            if(object == "scaled.regressed") {
                              
                                scaling <<- "_scale_nUMIregress"
                                pca.result <<- "pca.scaled.regressed"
                                object.path <<- path.scaled.regressed.object
                              
                            }
}


library(foreach)
library(doParallel)

# cores <- detectCores()
# cl <- makeCluster(2)
# doParallel::registerDoParallel(cl)

# foreach(object=names(seurat.objects)) %dopar% {

for (object in names(seurat.objects)) {
  
    print(object)
  
    selectObject(object)
  
    print(paste(object, pca.result, scaling, pca.path))
  
    pushbullet.start(message = paste("PCA for", object))
    
    tictoc::tic(paste("PCA for", object))

          assign(pca.result, doFastPCA(t(seurat.objects[[object]]@scale.data), 50,
                 center.use = F, scale.use = F, iterative = F)
          )


          saveRDS(pca.result, paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/pca/pcaObject.", age, scaling, ".Rds"))

      pushbullet.toc()
    
}

```


### 2.2 Project cells onto PCA space:

```{r}
pca.objects <- list(pca.scaled=pca.scaled, pca.scaled.regressed=pca.scaled.regressed)

for (object in names(seurat.objects)){
  
  print(object)
  
  selectObject(object)
  
  print(pca.result)
  
  pushbullet.start(message = paste("PCA projection for", pca.result))
  
    tictoc::tic(paste("PCA projection for", pca.result))
    
        ev <- pca.objects[[pca.result]]$sdev^2
      
        u <- max(which(pca.objects[[pca.result]]$sdev^2 > (sqrt(length(row.names(seurat.objects[[object]]@data)) / length(colnames(seurat.objects[[object]]@data))) + 1) ^2))
      
        sig_PCAs <- pca.objects[[pca.result]]$rotation[ , 1:u]
    
        cells_projected_sig_PCAs <- t(seurat.objects[[object]]@scale.data) %*% sig_PCAs
          
    pushbullet.toc()
    
  pushbullet.start(message = paste("Save PCA results to Seurat object:", pca.result))
  
    tictoc::tic(paste("Save PCA results to Seurat object:", pca.result))
        
          seurat.objects[[object]] <- SetDimReduction(seurat.objects[[object]], reduction.type = "fastPCA", slot = "cell.embeddings", 
                   new.data = cells_projected_sig_PCAs)
    
          seurat.objects[[object]] <- SetDimReduction(seurat.objects[[object]], reduction.type = "fastPCA", slot = "gene.loadings", 
          new.data = pca.objects[[pca.result]]$rotation)
    
          seurat.objects[[object]] <- SetDimReduction(seurat.objects[[object]], reduction.type = "fastPCA", slot = "sdev", 
          new.data =pca.objects[[pca.result]]$sdev)
        
          saveRDS(cells_projected_sig_PCAs, paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/pca/", age, pca.result, "projection.Rds"))
    
    pushbullet.toc() 

  
}
```

## 3. Clustering

### 3.1 Nearest neighbor
### 3.2 Jaccard distances
### 3.3 Graph-based clustering
### 3.4 Find cluster markers
```{r}

for (object in names(seurat.objects)) {
  
    selectObject(object)
  
    print(object)
  
    print(pca.result)
  
    print(object.path)
  
## 3.1 Nearest neighbor:
  
    pushbullet.start(message=paste("Nearest Neighbor", scaling))
    
        tic(paste("Nearest neighbor", scaling))
        
            cell.embeddings <- seurat.objects[[object]]@dr$fastPCA@cell.embeddings
            
            nearest <- nn2(cell.embeddings, k=10)
            
            rownames(nearest$nn.idx) <- rownames(cell.embeddings)
            
            
            out.nn.idx <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/nn2_output_neighbors_sigPCAs_", age, scaling, ".txt")
            write.table(nearest$nn.idx, file=out.nn.idx, sep="\t", quote=F, col.names=NA)
            
            out.nn.dists <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/nn2_output_distance_sigPCAs_", age, scaling,".txt")
            write.table(nearest$nn.dists, file=out.nn.dists, sep="\t", quote=F, col.names=NA)
        
        pushbullet.toc()
    
### 3.2 Jaccard distance:
    
  pushbullet.start(message=paste("Jaccard distances", scaling))


        tic(paste("Jaccard distances", scaling))
          
            jaccard.out <- paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/jaccard_weighted_edges_", age, scaling, "_fromR.txt")
            
            system(paste0("perl /kriegsteinlab/data1/carmen/2nd-trimester/clustering/code/jaccard_toedges2.pl ", out.nn.idx, " ", out.nn.dists, " > ", jaccard.out))
            
            edgedata <- read.table(jaccard.out, sep="\t", header=T)
    
            edges <- as.data.frame(edgedata)
    
            pushbullet.toc()
    
### 3.3 Graph-based clustering:
        
  pushbullet.start(message=paste("Graph-based Clustering", scaling))
          
      tic(paste("Calculate clusters", scaling))
      
              input <- graph_from_data_frame(edges, directed=FALSE)
      
              clusters <- cluster_louvain(input)
      
              clustered <- membership(clusters)
      
              names.col1.file <- paste0("names.in.col1_", age, scaling, ".txt")
      
              write(clustered, file=names.col1.file, ncol=1)
      
      pushbullet.toc()


names <- edgedata[ , 1]

uniquenames <- unique(names)

meta <- read.table(file=names.col1.file, sep="\t")

rownames(meta) <- uniquenames

seurat.objects[[object]] <- AddMetaData(seurat.objects[[object]], meta, "groups")

seurat.objects[[object]] <- SetAllIdent(seurat.objects[[object]], id="V1")

sizes(clusters)

write.table(seurat.objects[[object]]@ident, paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/clusterIdentity/clusterIdentity_", age, "_", scaling, ".txt"), sep="\t", quote=F, col.names=NA)


### 3.4 Find cluster markers:

pushbullet.start(message=paste("Find cluster markers for", scaling))
          
      tic(paste("Find cluster markers for", scaling))
      

cluster.markers <- FindAllMarkers(seurat.objects[[object]], min.pct = 0.25, thresh.use = 0.25, only.pos=TRUE, max.cells.per.ident=2000)

write.table(cluster.markers, paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/clustermarkers_", age, "_", scaling, ".txt"), sep="\t", quote=F, col.names=T)

}

pushbullet.toc()

saveRDS(seurat.objects[[object]], file=object.path)
```


## 4. Run tSNE:
```{r}

for (object in c(seurat.obj.scale, seurat.obj.scale.regress)) {
  
  selectObject()
  
    tic(paste("Run tSNE for", age, scaling))


        object <- RunTSNE(object, reduction.use = "fastPCA", perplexity=30, dims.use = 1:ncol(sig_PCAs), do.fast = F, check_duplicates=FALSE)
        
        saveRDS(object@dr$tsne@cell.embeddings, file = paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/tsne/tsne_", age, "_", scaling, ".txt"))
        
        tic(paste("Plot TSNE", age))

            pdf(paste0("/kriegsteinlab/data1/carmen/2nd-trimester/clustering/tsne", scaling, "_", perplexity,  "_", age, ".pdf"))
    
               TSNEPlot(object, do.label=TRUE)
    
            dev.off()

        pushbullet.toc()
  
        saveRDS(object, file = object.path)
        
        
    pushbullet.toc()

}

# Rtsne:

# Rtsne(gw22@dr$fastPCA@cell.embeddings, dims = 2, initial_dims = 7, perplexity = 5,
# theta = 0, check_duplicates = TRUE, pca = FALSE, max_iter = 1000,
# verbose = TRUE, is_distance = FALSE, Y_init = NULL)
```


