---
title: 10X Standard QC Plots
author: Carmen
date: Aug 30, 2018
output:
  html_document:
    df_print: paged
---

### User input: Prepare session.

To run this R Markdown file, enter the following lines in R:
```{r eval=FALSE}
library(rmarkdown)
library(knitr)
```

Age of the dataset. 
```{r eval=FALSE}
age <- "GW14"
```

Choose an upper limit for nGenes and nUMIs for zoomed-in plots.
```{r eval=FALSE}
nGene.limit <- 3500
nUMI.limit <- 10000
```

Indicate the path to your list of ribosomal gene names.
```{r eval=FALSE}
ribogenes <- read_table("/kriegsteinlab/data1/carmen/2nd-trimester/qc/ribogenes.txt", col_names = F)$X1
```

Path to your dataset (in this case, an `.RData` workspace file), and the name of the Seurat object contained within the workspace.
We load the data before running the `.Rmd` file as this takes a long time, and is otherwise done everytime the script is run (immune to caching.)
```{r eval=FALSE}
dataset <- "/kriegsteinlab/data1/aparna/homefiles/gw14.RData"
load(dataset)

seurat.obj <- "gw14"
```

Indicate the path to a metadata file with cell names and their brain regions of origin.
```{r eval=FALSE}
meta <- read_tsv("/kriegsteinlab/data1/carmen/2nd-trimester/metadata/colnames_GW14_080818_fixed.txt", 
                   col_names = c("cell.name", "brain.region"))
```

Call the `knitr::render` function indicating the path to this `.Rmd` file:
```{r eval=FALSE}
render("/kriegsteinlab/data1/carmen/2nd-trimester/qc/code/QC_10X_plots_2018-08-30.Rmd")
```

### Begin R Markdown script:
```{r knitr-opts_chunk}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, cache.lazy = FALSE)
```

### Load libraries
```{r load-libraries, cache=TRUE}
library(Seurat)
library(readr)
library(ggplot2)
library(scales)
library(viridis)
library(tidyr)
library(dplyr)
library(janitor)
```

### Create a new Seurat object using the raw data in the original Seurat object.
For compatibility issues with the new version of Seurat.
```{r import-data, cache=TRUE}
assign("seurat.obj.new", CreateSeuratObject(raw.data = get(seurat.obj)@raw.data, min.cells = 0, min.genes = 0, project=age))

rm(list=seurat.obj)
```

### Declare multiplot function.
```{r multiplot-fn, cache = TRUE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

### Set the brain region of cells as their original identity in both the `@ident` and the `@meta.data` slots.
```{r set-ident, cache=TRUE}
seurat.obj.new@meta.data$orig.ident <- factor(meta$brain.region)

seurat.obj.new <- SetIdent(seurat.obj.new, cells.use = WhichCells(seurat.obj.new), ident.use = meta$brain.region)
```

```{r meta-summary, cache=TRUE}
summary(seurat.obj.new@meta.data)
```

### Generate table with number and percentage of cells in each brain region.
```{r brainregions-table, cache=TRUE}
seurat.obj.new@meta.data %>% tabyl(orig.ident) %>% adorn_totals(where="row") %>% arrange(desc(n)) 
```

### Seurat's ViolinPlot plotting by area.
(Not working currently.)
```{r seurat-violin, eval=FALSE}
violin.plot <- VlnPlot(seurat.obj.new, features.plot = c("nGene", "nUMI"), do.sort = T, point.size.use = .5, ident.include = as.character(statsByArea$orig.ident))
```

### Calculate % of mitochondrial genes in each cell.
```{r calc-mito, cache=TRUE}
mito.genes <- grep(pattern="^MT-", x=rownames(seurat.obj.new@data), value = TRUE)
percent.mito <- colSums(seurat.obj.new@raw.data[mito.genes, ])/colSums(seurat.obj.new@raw.data)

seurat.obj.new <- AddMetaData(object = seurat.obj.new, metadata = percent.mito, col.name = "percent.mito")
```

### Calculate % of ribosomal genes in each cell.
```{r calc-ribo, cache=TRUE}
ribo <- seurat.obj.new@raw.data[ribogenes, ]
percent.ribo <- colSums(ribo)/colSums(seurat.obj.new@raw.data)

seurat.obj.new <- AddMetaData(object = seurat.obj.new, metadata = percent.ribo, col.name = "percent.ribo")
```

### Calculate mean UMIs, genes, ribosomal and mitochondrial % for each brain region.
```{r calc-stats-by-area, cache=TRUE}
statsByArea <- seurat.obj.new@meta.data %>% dplyr::group_by(orig.ident) %>% 
  summarise(mean.nGene = mean(nGene), mean.nUMI =mean(nUMI), mean.pctmito = mean(percent.mito), mean.pctribo= mean(percent.ribo))

statsByArea
```


## PLOTS

### 1. Violin plot: Genes per cell, split by brain region.
```{r p1, cache=TRUE}
seurat.obj.new@meta.data$orig.ident <- ordered(seurat.obj.new@meta.data$orig.ident, 
                                       levels = (statsByArea %>% arrange(mean.nGene))$orig.ident)

p1 <- ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=orig.ident), alpha=0.2, size=0.25) + 
  geom_violin(aes(orig.ident, nGene), colour="grey60", scale = "width", alpha=0.3) +
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  scale_y_continuous(breaks=seq(0, max(seurat.obj.new@meta.data$nGene), by=1000)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Genes per Cell"),
       x="", 
       y="Genes per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p1
```


### 2. Violin plot: Genes per cell [zoom-in].
```{r p2, cache=TRUE}
p2 <- ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=orig.ident), alpha=0.2, size=0.25) + 
  geom_violin(aes(orig.ident, nGene), colour="grey60", scale = "width", alpha=0.3) +
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  scale_y_continuous(breaks=seq(0, nGene.limit, by=nGene.limit/10), limits = c(0, nGene.limit)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Genes per Cell (Zoom)"),
       x="", 
       y="Genes per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)
  
p2
```



### 3. Violin plot: UMIs per cell, split by brain region.
```{r p3, cache=TRUE}
p3 <- 
  ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nUMI, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60", colour="grey30", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  scale_y_continuous(breaks=seq(0, max(seurat.obj.new@meta.data$nUMI), by=10000)) +
  labs(title=paste0(age, ": UMIs per Cell"),
       x="", 
       y="UMIs per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p3
```

### 4. Violin plot: UMIs per cell [Zoom-in].
```{r p4, cache=TRUE}
p4 <- 
  ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nUMI, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60",  colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, nUMI.limit, by=nUMI.limit/10), limits=c(0,  nUMI.limit)) +
  labs(title=paste0(age, ": UMIs per Cell (Zoom)"),
       x="", 
       y="UMIs per Cell"
      ) +
  guides(fill=FALSE, colour=FALSE)

p4
```

### 5. Violin plot: Genes per cell + color by nUMI.
```{r p5, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(nUMI)

p5 <- ggplot(seurat.obj.new@meta.data %>% filter(nGene < nGene.limit)) + 
  geom_jitter(aes(orig.ident, nGene, colour=log10(nUMI)), alpha=0.6, size=0.25) + 
  scale_color_viridis(option="magma", begin = .1, end = .95) +
  geom_violin(aes(orig.ident, nGene), colour="grey60",  colour="grey60", alpha=0.1, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.9, 1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 6),
        axis.text.x = element_text(angle = 30)
        # legend.title.align = 0.9
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, nGene.limit, by=nGene.limit/10), limits=c(0, nGene.limit)) +
  labs(title=paste0(age, ": Genes per Cell / log UMIs per Cell"),
       x="", 
       y="Genes Detected",
       color= "log10 UMI")

p5
```

### 6. Violin plot: UMIs per cell + color by nGenes.
```{r p6, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(nGene)

p6 <- ggplot(seurat.obj.new@meta.data %>% filter(nUMI < nUMI.limit)) + 
  geom_jitter(aes(orig.ident, nUMI, colour=nGene), alpha=0.75, size=0.25) + 
  scale_color_viridis(option="magma", begin = .1, end = .9) +
  geom_violin(aes(orig.ident, nUMI), colour="grey60", alpha=0.1, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.9,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 12),
        axis.text.x = element_text(angle = 30)
        # legend.title.align = 0.9
        ) +
  # ylim(0, 40000) +
  scale_y_continuous(breaks=seq(0, nUMI.limit, by=nUMI.limit/10), limits=c(0, nUMI.limit)) +
  labs(title=paste0(age, ": UMIs per Cell / Genes per Cell"),
       x="", 
       y="UMIs Detected",
       color= "")

p6
```

### 7. % Ribosomal, split by brain region.
```{r p7, cache=TRUE}
p7 <- 
ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, percent.ribo, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  geom_violin(aes(orig.ident, percent.ribo), colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust=0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  # scale_y_continuous(breaks=seq(0,40000, by=5000)) +
  # ylim(0,40000) +
  labs(title=paste0(age, ": % Ribosomal Genes per Cell"),
       x="", 
       y="% Ribosomal Genes"
      ) +
  guides(fill=FALSE, colour=FALSE)

p7
```

### 8. % Mitochondrial, split by brain region.
```{r p8, cache=TRUE}
p8 <- 
  ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, percent.mito, colour=orig.ident), alpha=0.2, size=0.25) + 
  scale_color_manual(values = c("#E58606", "#5D69B1", "#52BCA3", "#99C945", "#ffc300" ,"#ED645A", "#2F8AC4", "#764E9F", "#E58606", "#5D69B1", "#52BCA3", "#99C945")) +
  geom_violin(aes(orig.ident, percent.mito), colour="grey60", alpha=0.3, scale="width") +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust=0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  # scale_y_continuous(breaks=seq(0,0.1, by=0.01), limits = c(0, 0.1)) +
  
  labs(title=paste0(age, ": % Mitochondrial Genes per Cell"),
       x="", 
       y="% Mitochondrial Genes"
      ) +
  guides(fill=FALSE, colour=FALSE)

p8
```

### 9. Violin plot: Genes per cell; color by % ribosomal genes.
```{r p9, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.ribo)

ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.ribo), alpha=0.6, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey40", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .9) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  scale_y_continuous(breaks=seq(0, max(seurat.obj.new@meta.data$nGene), by=1000)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Genes Detected per Cell"),
       x="Area", 
       y="nGene"
      )
```

### 10. Violin plot: Genes per cell; color by % ribosomal genes [zoom-in]
```{r p10, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.ribo)

p10 <- ggplot(seurat.obj.new@meta.data %>% filter(nGene < nGene.limit)) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.ribo), alpha=0.8, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey40", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .1, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 30)
        # legend.title.align = 0.9
        ) +
  scale_y_continuous(breaks=seq(0, nGene.limit, by=nGene.limit/10), limits = c(0, nGene.limit)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Genes per Cell / % Ribosomal Genes\n(zoom-in)"),
       x="", 
       y="Genes per Cell",
       color="% ribosomal"
      )

p10
```


### 11. Violin Plot: Genes per cell, color by % mitochondrial genes
```{r p11, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.mito)

p11 <- ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.mito), alpha=0.7, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .9) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        axis.text.x = element_text(angle = 30)
        ) +
  scale_y_continuous(breaks=seq(0, max(seurat.obj.new@meta.data$nGene), by=500)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Number of Genes Detected per Cell"),
       x="Area", 
       y="nGene"
      )

p11
```

### 12. Violin Plot: Genes per cell, color by % mitochondrial genes [zoom-in]
```{r p12, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.mito)

p12 <- ggplot(seurat.obj.new@meta.data %>% filter(nGene < nGene.limit)) + 
  geom_jitter(aes(orig.ident, nGene, colour=percent.mito), alpha=0.8, size=0.25) +
  geom_violin(aes(orig.ident, nGene), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .15, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,1),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        # legend.title.align = 0.9
        ) +
  scale_y_continuous(breaks=seq(0, nGene.limit, by=nGene.limit/10), limits = c(0, nGene.limit)) +
  scale_x_discrete(breaks=(statsByArea %>% arrange(mean.nGene))$orig.ident) +
  labs(title=paste0(age, ": Genes per Cell / % Mitochondrial Genes"),
       x="", 
       y="Genes per Cell",
       color="% mito"
      )

p12
```

### 13. Violin Plot: % Ribosomal, color by % mitochondrial.
```{r p13, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.mito)

p13 <- 
ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, percent.ribo, colour=percent.mito), alpha=0.7, size=0.25) + 
  geom_violin(aes(orig.ident, percent.ribo), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .10, end = .90) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.87,.95),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 30)
        # legend.title.align = 0.9
        ) +
  labs(title=paste0(age, ": % Ribosomal Genes / % Mitochondrial Genes per Cell"),
       x="", 
       y="% Ribosomal Genes",
       color="% mito"
      )

p13
```


### 14. Violin Plot: % Mitochondrial, color by % ribosomal.
```{r p14, cache=TRUE}
seurat.obj.new@meta.data <- seurat.obj.new@meta.data %>% arrange(percent.ribo)

p14 <- 
ggplot(seurat.obj.new@meta.data) + 
  geom_jitter(aes(orig.ident, percent.mito, colour=percent.ribo), alpha=0.7, size=0.20) + 
  geom_violin(aes(orig.ident, percent.mito), colour="grey60", alpha=0.1, scale="width") +
  scale_color_viridis(option="magma", begin = .10, end = .95) +
  theme_minimal() +
  theme(plot.title = element_text(size= 12, hjust = 0.5), 
        axis.title.y=element_text(vjust = 2),
        axis.title.x=element_text(vjust = -1),
        legend.position=c(0.85,.93),
        legend.key.size = unit(0.5,"cm"),
        legend.direction = "horizontal",
        legend.text = element_text(angle = 30, size = 7),
        legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 30)
        # legend.title.align = 0.9
        ) +
  labs(title=paste0(age, ": % Mitochondrial Genes / % Ribosomal Genes per Cell"),
       x="", 
       y="% Mitochondrial Genes",
       color = "% ribosomal"
      )

p14
```

```{r remove-large-objects}
meta.from.seurat <- seurat.obj.new@meta.data
rm(seurat.obj.new, matrix_scaled, matrix_pca, ribo, cluster)
```

### Print all QC plots in a single .png file with high res for easy sharing.
```{r print-plots}
png(paste0(age, "_QC_all.png"), width=25, height=16, units="in", res=120)

multiplot(p1, p3, p7, p8, p2, p4, p10, p12, p5, p6, p13, p14, cols=3)

dev.off()
```

